Finally tagless, partially evaluated

### 2007-11-25 tagless/talk.tex
Some more slides with not much content

\begin{frame}{Simple type preservation}
    \def\bool{\only<5-6>{\ \alert{bool}}}
    \begin{tikzpicture}
        \draw node [lang] (lang)
          {\texttt{\alt<6->{\onslide<6>{\alert{leq} (\alert{lit} 3, }%
                            \only<7>{\makebox[0pt]{\sffamily
                                      Simply typed $\lambda$-calculus}}%
                            \onslide<6>{\alert{lit} 4):}}
                           {\onslide<2->{LEQ (LIT 3, }%
                            \only<1>{\makebox[0pt]{\sffamily
                                      \alert{Typed} source language}}%
                            \onslide<2->{LIT 4):}}%
                   \makebox[3em][l]{\bool\only<2-5>{\ term}}}};
        \draw [->] (lang) --
            node [above,midway,anchor=east]
              {\only<1>{\alert{Typed} metalanguage\hspace*{2em}}%
               \only<2>{evaluate}%
               \only<7>{\begin{tabular}{@{}r@{\qquad}}
                        Haskell constructor classes\\
                        or ML functors
                        \end{tabular}}}
            node [above,midway,anchor=east,text width=2.5in,inner xsep=0pt]
                {\only<-2,7>{\color{bg}}\ttfamily
                \alt<6->{\only<6>{\alert<6>{lit:\ int -> int}\\
                lit i = i\\[1ex]
                \alert<6>{leq:\ int * int -> bool}\\
                leq (i,j) = i <= j}}
                {\only<2-5>{\alert<5>{eval:\ \only<5>{'a }term -> \only<5>{'a }value}\\
                eval (LIT i) = INT i\\
                eval (LEQ (m,n)) =\\
                \quad \alert<4>{match (eval m, eval n) with\\
                \quad (INT i, INT j) ->} BOOL (i <= j)}}}
            +(-100:4cm)
            node [anchor=north,text depth=0pt,text height=9pt]
              {\texttt{\llap{\onslide<2-5>{BOOL}
                             \onslide<2-6>{true:}%
                             \only<1>{\rlap{\sffamily
                                            \alert{Typed} target languages}}%
                             \only<7>{\rlap{\sffamily evaluate}}}%
                       \makebox[3em][l]{\bool\only<2-5>{\ value}}}};
        \draw [->] (lang) -- +(-80:3.5cm)
            node [anchor=north west,inner xsep=0pt] {\only<7>{\hspace*{-16pt}compile}};
        \draw [->] (lang) -- +(-65:3cm)
            node [anchor=north west,inner xsep=0pt] {\only<7>{specialize}};
        \draw [->] (lang) -- +(-50:2.5cm)
            node [anchor=north west,inner xsep=0pt] {\only<7>{\hspace*{-3pt}CPS transform}};
        \draw      (lang)    +(-35:2cm) node {\dots};
    \end{tikzpicture}

@@

    \strut
    \only<-3>{It should be obvious in the metalanguage that interpreting
    a well-typed source term yields a well-typed target term.}%
    \only<4>{The term should be well-typed (and closed), so pattern matching
    in the metalanguage should always \textbf{obviously} succeed.}%
    \only<5>{Previous solutions use (and motivate) fancier types:\\
    generalized abstract data types, dependent types.}%
    \only<6>{Our simple solution is to be \textbf{finally tagless:}\\
    replace term constructors by cogen functions.}%
    \only<7>{The term accommodates \textbf{multiple interpretations}\\
    by abstracting over the cogen functions and their types.}%
    \strut
\end{frame}

>>>>

\begin{frame}{Simple type preservation}
    \def\bool{\only<6-7>{\ \alert{bool}}}
    \begin{tikzpicture}
        \draw node [lang] (lang)
          {\texttt{\alt<7->{\onslide<7>{\alert{leq} (\alert{lit} 3, }%
                            \only<8>{\makebox[0pt]{\sffamily
                                      Simply typed $\lambda$-calculus}}%
                            \onslide<7>{\alert{lit} 4):}}
                           {\onslide<3->{LEQ (LIT 3, }%
                            \only<1>{\makebox[0pt]{\sffamily
                                      \alert{Typed} source language}}%
                            \only<2>{\makebox[0pt]{$3 \le 4$}}%
                            \onslide<3->{LIT 4):}}%
                   \makebox[3em][l]{\bool\only<3-6>{\ term}}}};
        \draw [->] (lang) --
            node [above,midway,anchor=east]
              {\only<1>{\alert{Typed} metalanguage\hspace*{2em}}%
               \only<2-3>{evaluate}%
               \only<8>{\begin{tabular}{@{}r@{\qquad}}
                        Haskell constructor instances\\
                        or ML functors
                        \end{tabular}}}
            node [above,midway,anchor=east,text width=2.5in,inner xsep=0pt]
                {\only<-3,8>{\color{bg}}\ttfamily
                \alt<7->{\only<7>{\alert<7>{lit:\ int -> int}\\
                lit i = i\\[1ex]
                \alert<7>{leq:\ int * int -> bool}\\
                leq (i,j) = i <= j}}
                {\only<3-6>{\alert<6>{eval:\ \only<6>{'a }term -> \only<6>{'a }value}\\
                eval (LIT i) = INT i\\
                eval (LEQ (m,n)) =\\
                \quad \alert<5>{match (eval m, eval n) with\\
                \quad (INT i, INT j) ->} BOOL (i <= j)}}}
            +(-100:4cm)
            node [anchor=north,text depth=0pt,text height=9pt]
              {\texttt{\llap{\onslide<3-6>{BOOL}
                             \onslide<3-7>{true:}%
                             \only<1>{\rlap{\sffamily
                                            \alert{Typed} target languages}}%
                             \only<2>{\rlap{\enspace$\mathsf{true}$}}%
                             \only<8>{\rlap{\sffamily evaluate}}}%
                       \makebox[3em][l]{\bool\only<3-6>{\ value}}}};
        \draw [->] (lang) -- +(-80:3.5cm)
            node [anchor=north west,inner xsep=0pt] {\only<8>{\hspace*{-16pt}compile}};
        \draw [->] (lang) -- +(-65:3cm)
            node [anchor=north west,inner xsep=0pt] {\only<8>{specialize}};
        \draw [->] (lang) -- +(-50:2.5cm)
            node [anchor=north west,inner xsep=0pt] {\only<8>{\hspace*{-3pt}CPS transform}};
        \draw      (lang)    +(-35:2cm) node {\dots};
    \end{tikzpicture}

@@

    \strut
    \only<-4>{It should be obvious in the metalanguage that interpreting
    a well-typed source term yields a well-typed target term.}%
    \only<5>{The term should be well-typed (and closed), so pattern matching
    in the metalanguage should always \textbf{obviously} succeed.}%
    \only<6>{Previous solutions use (and motivate) fancier types:\\
    generalized abstract data types, dependent types.}%
    \only<7>{Our simple solution is to be \textbf{finally tagless:}\\
    replace term constructors by cogen functions.}%
    \only<8>{The term accommodates \textbf{multiple interpretations}\\
    by abstracting over the cogen functions and their types.}%
    \strut
\end{frame}

\section{The object language}
\subsection{As a constructor class in Haskell}
\subsection{As a functor signature in ML}
\section{Tagless interpretation}
\subsection{Evaluation}
\subsection{Compilation}
\section{Type-indexed types}
\subsection{Partial evaluation}
\subsection{CPS transformation}

\begin{frame}{Conclusion}
    Initial type-checking can be done

    Preserves sharing
\end{frame}
